---
title: "2015Sichuan-MBC"
output: html_document
---

Comparing two dataframes in R
http://stackoverflow.com/questions/30655456/comparing-data-frames-in-r/30655765

Set up github remote on github.com/dougwyu/Sichuan2014
https://www.r-bloggers.com/rstudio-and-github/
Open Terminal using Tools...Shell..
```{bash}
git remote add origin https://github.com/dougwyu/Sichuan2015.git
git config remote.origin.url git@github.com:dougwyu/Sichuan2015.git
git pull -u origin master
git push -u origin master
```
Pull and Push buttons are now active


```{r setup, test methods}
# 2libs_otu_WithTax4R_orig.txt is original file with all taxonomic assignments. Some of the OTUs still have non-Arthropoda in them, so need to go back and make taxonomic subsets. I added "Site" to top of first column and fixed misspelling of kindom to kingdom
##### make sure that Source option is set to `Chunk output in Console` and do not select `Show Previews inline`
# before import to R, delete first line and "#OTU ID" in the second line by hand. change names of habitats type: Ag->CL, B->BB, E->EC, C->JC, M->MF. add habitats type to original table. separate taxon info by "\t"

# add packages that I will use
library(tidyverse) # includes all data-formatting packages as one big package (e.g. dplyr, tidyr, ggplot2, readr, readxl, tibble, and others)
library(broom)
library(vegan)
library(car)
library(breakaway)
library(boral)
library(mvabund)
```

```{r load and format data}
# help(read_tsv)
# this inputfile contains taxononomic assignments in the lower rows
# command from readr package, and you tell the command how to format the columns
inputfile <- "table_LerayCOI_96.txt"
habitatinput <- "habitat.txt"

gfgMB <- read_tsv(inputfile, col_names = TRUE, na = "NA")
gfgMB <- tbl_df(gfgMB)

habitat <- read_tsv(
	habitatinput, col_names = TRUE, na = "NA", 
	col_types = cols(
	  PlotID = col_character(),
	  Region = col_character(),
	  Location = col_character(),
	  TreeCover = col_integer(),
	  Clumpedness = col_integer()
		)
)
```

```{r community dataset}
# make community dataset
community <- gfgMB
community <- community %>% select(-pos) # remove positive control
community_t <- t(community)
community_t <- as.data.frame(community_t)
colvector <- t(community[, 1])
colnames(community_t) <- colvector # add column names
community_t <- community_t[-1, ]
# community_t <- rownames_to_column(community_t)

# convert the columns to numeric from factor
# http://stackoverflow.com/questions/2288485/how-to-convert-a-data-frame-column-to-numeric-type
community_t <- sapply(community_t, function(x) as.numeric(as.character(x))) # sapply applies a function to each column, and the function is:  function(x) as.numeric(as.character(x)).  Cannot convert factors to numeric directly. first convert to character, then to numeric
community_t <- as.data.frame(community_t) # then convert to df
colSums(community_t) # some colSums = 0 because these are the positive control OTUs
```

```{r remove pos ctrl OTUs}
community_t <- community_t[ ,which(colSums(community_t) > 0)]
colSums(community_t)
rowSums(community_t)
```


```{r choose commmunity to analyse}
community_name <- deparse(substitute(community_t)) # put name of community here

habitat_now <- habitat # 

community_now <- get(community_name)
```

```{r no_singletons}
community_now_nosngltons <- community_now %>% select_if(specnumber(community_now, MARGIN = 2) > 1) # remove singleton OTUs (found in only one site) specnumber(MARGIN = 2) counts the number of sites that have a particular OTU. (same as colSums > 2)
rowSums(community_now_nosngltons) #  rowSums = 0 in row 53
# remove any rows (sites) with no species
	# first use the community table to alter the habitat table
habitat_now_nosngltons <- habitat_now[which(rowSums(community_now_nosngltons) > 0), ]
	# then alter the community tables
community_now_nosngltons <- community_now_nosngltons[which(rowSums(community_now_nosngltons) >0), ]
rowSums(community_now_nosngltons)
```

visualise a histogram of read number per OTU
```{r histogram of otu read numbers}
##### calculate distribution of read numbers per OTU to set minimum number 
otureads <- c(colSums(community_now)) # list of the reads per OTU
sum(otureads) ## 1,086,199 reads total (this is after removing some of the sites, so this is almost but not all the reads in the original dataset)
otureads[otureads>5000] <- 5000 # to make the histogram readable
otuhist <- hist(otureads, breaks = 100)
text(otuhist$mids, otuhist$counts, cex = 0.5, otuhist$counts, adj = c(.5, -.5), col = "blue3")
```


```{r}
# vegemite(community_now)
dca <- decorana(community_now)
vegemite(community_now, dca, "Hill")

plotree <- hclust(vegdist(community_now), "average")
tabasco(community_now, plotree)
```

```{r nmds}
# do NMDS analysis to see basic patterns #
community.jmds <- metaMDS(community_now, distance = "jaccard", trymax = 40, binary = FALSE)
community.jmds <- metaMDS(community_now, distance = "jaccard", binary = FALSE, previous.best = community.jmds)

stressplot(community.jmds)
```

```{r plot nmds}
#### plot the communities
(sprichness <- specnumber(community_now, MARGIN = 1)) # number of species per site

with(habitat_now, ordisurf(community.jmds, sprichness, main=community_name, cex=Clumpedness*TreeCover/20))

with(habitat_now, ordisurf(community.jmds, TreeCover, main=community_name, cex=TreeCover/5))

with(habitat_now, ordisurf(community.jmds, Clumpedness, main=community_name, cex=0.5))

with(habitat_now, ordispider(community.jmds, Region, main=community_name, cex=0.5))

with(habitat_now, ordiellipse(community.jmds, Region, cex=.5, draw="polygon", col=c("blue"), alpha=20, kind="se", conf=0.95, label=TRUE))
# with(habitat_now, ordisurf(community.jmds, Altitude, main="", cex=0.5))

```

```{r ordiplots, eval = FALSE}
# These commands are ways to identify points
orditorp(community.jmds, labels = habitat_now$Site, dis="sites", pcol = "gray")

ordipointlabel(community.jmds, dis="sites")

p1 <- plot(community.jmds, dis="sites")
identify(p1, "sites") # interactive method to identify sites by row number. Click on the points that you want to identify. When you have clicked on all the points, click on the Finish button at top right of Plot window. 
```



```{r boral, eval=FALSE}
#mcmc.control <- list(n.burnin = 100, n.iteration = 1000, n.thin = 10)
mcmc.control <- list(n.burnin = 300, n.iteration = 3000, n.thin = 30)
colnames(community_now) <- c(1:645) # change colnames to a simple number
community_now[community_now > 1] <-  1 # change to presence/absence
comm.fit.p3 <- boral(community_now, family = "poisson", num.lv = 3, row.eff = "fixed", calc.ics = FALSE, mcmc.control = mcmc.control)
summary(comm.fit.p3)
par(mfrow = c(2,2))
plot(comm.fit.p3) ## Plots used in residual analysis, 
par(mfrow = c(1,1))
lvsplot(comm.fit.p3, biplot = FALSE, jitter = TRUE)

```



``` {r}

# otureads200max <- otureads[otureads>1] # remove otus with 1 read  # has no effect if I have removed singleton reads already
otureads[otureads>100] <- 100 # replace large read numbers with 100, to make the histogram readable
hist(otureads)  # bimodal distribution of read numbers. Now, what should be the cutoff point? 

commNosr.100 <- community_now[ ,colSums(community_now)>=100] ## remain 346 otus, community without Non-single reads otu
commNosr.080 <- community_now[,colSums(community_now)>=80]
commNosr.070 <- community_now[,colSums(community_now)>=70]
commNosr.060 <- community_now[,colSums(community_now)>=60]
commNosr.050 <- community_now[,colSums(community_now)>=50]  
commNosr.030 <- community_now[,colSums(community_now)>=30]
commNosr.020 <- community_now[,colSums(community_now)>=20] 
commNosr.010 <- community_now[,colSums(community_now)>=10]
commNosr.002 <- community_now[,colSums(community_now)>=2] # this removes single-read OTUs (an OTU's number of reads is its colSum)

commNosr <- commNosr.100

commNosr.jmds <- metaMDS(commNosr, distance = "jaccard", trymax = 40)
commNosr.jmds <- metaMDS(commNosr, distance = "jaccard", previous.best = commNosr.jmds)
stressplot(commNosr.jmds)

community.jmds <- commNosr.jmds

levels(habitat_now$Habitat)
with(habitat_now, ordisurf(community.jmds, Altitude, main="", cex=0.5))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("blue"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BB"))))
# with(habitat_now, ordispider(community.jmds, Habitat, main="", show.groups=(c("BB"))))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("green"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("CL"))))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("red"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("EC"))))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("blue"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("JC"))))
# with(habitat_now, ordispider(community.jmds, Habitat, main="", show.groups=(c("JC"))))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("darkgreen"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("MF"))))
with(habitat_now, ordispider(community.jmds, Habitat, main="", show.groups=(c("MF"))))

with(habitat_now, ordiellipse(community.jmds, Habitat, cex=.5, draw="polygon", col=c("darkgreen"), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("NF"))))
with(habitat_now, ordispider(community.jmds, Habitat, main="", show.groups=(c("NF"))))
```

http://www.wcsmalaysia.org/analysis/Biod_richness.htm
```{r t.test2}
############ This function (t.test2) will calculate Welch's test
t.test2 <- function(m1, m2, s1, s2, n1, n2, m0=0, equal.variance=FALSE)
{
  if( equal.variance==FALSE ) 
  {
    se <- sqrt( (s1^2/n1) + (s2^2/n2) )
    # welch-satterthwaite df
    df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
  } else
  {
    # pooled standard deviation, scaled by the sample sizes
    se <- sqrt( (1/n1 + 1/n2) * ((n1-1)*s1^2 + (n2-1)*s2^2)/(n1+n2-2) ) 
    df <- n1+n2-2
  }      
  t <- (m1-m2-m0)/se 
  dat <- c(m1-m2, se, t, round(df,1), 2*pt(-abs(t),df))    
  names(dat) <- c("Difference of means", "Std Error", "t", "df", "p-value")
  return(dat) 
}
```

```{r specpool}
# IMPORTANT:  Don't use OTU tables with singleton OTUs removed, because singleton OTUs are needed for chao stats

# using community_insecta
pool1 <- specpool(community_now)
pool1

# chao, MF vs NF
# t.test2(3513.490, 9671.864, 543.1159*sqrt(9), 981.5702*sqrt(15), 9, 15)
t.test2(pool1[5, 3], pool1[6, 3], pool1[5, 4]*sqrt(pool1[5, 10]), pool1[6, 4]*sqrt(pool1[6, 10]), pool1[5, 10], pool1[6, 10])
# jack1, MF vs NF
# t.test2(1013.4444, 2603.9333, 190.4298*sqrt(9), 477.7467*sqrt(15), 9, 15)
t.test2(pool1[5, 5], pool1[6, 5], pool1[5, 6]*sqrt(pool1[5, 10]), pool1[6, 6]*sqrt(pool1[6, 10]), pool1[5, 10], pool1[6, 10])


# using commNosr.n (OTUs with >= n total reads only)
pool2 <- specpool(commNosr.010, habitat_now$Habitat) %>% rownames_to_column("Habitat")
pool2

# chao, MF vs NF
# t.test2(188.19048, 313.98718, 31.812022*sqrt(9), 47.707941*sqrt(15), 9, 15) # commNosr.100
t.test2(pool2[5, 3], pool2[6, 3], pool2[5, 4]*sqrt(pool2[5, 10]), pool2[6, 4]*sqrt(pool2[6, 10]), pool2[5, 10], pool2[6, 10])
# jack1, MF vs NF
# t.test2(154.6667, 240.6667, 24.741628*sqrt(9), 30.652170*sqrt(15), 9, 15) # commNosr.100
t.test2(pool1[5, 5], pool2[6, 5], pool2[5, 6]*sqrt(pool2[5, 10]), pool2[6, 6]*sqrt(pool2[6, 10]), pool2[5, 10], pool2[6, 10])

```

The more low-read-number OTUs are included in the dataset, the more that the species richness of NF > MF. This might imply that the diversity of NF is in the low biomass species or that, for some reason, NF has more OTU artefacts

```{r shannon diversity}
community_insecta %>% filter(habitat$Habitat == "MF") %>% diversity(index = "simpson") %>% mean()
community_insecta %>% filter(habitat$Habitat == "NF") %>% diversity(index = "simpson") %>% mean()
```

#### mvabund analysis

```{r mvabund}
community_now_mva<-mvabund(community_now) # convert the arthropods dataset to mvabund format
is.mvabund(community_now_mva) # checking that the arthmvabund dataset is properly formatted
meanvar.plot(community_now_mva)
plot(community_now_mva ~ habitat_now$Habitat) # predictor must be a factor ~ habitat_now$habitat) 
```

```{r mvabund model}
comm.nb <- manyglm(community_now_mva ~ habitat_now$Habitat, family="negative.binomial") # alternatives are:  binomial, binomial(link="cloglog"), poisson
```

```{r residuals pit.norm}
plot(comm.nb, res.type="pit.norm") # residuals vs fitted, pretty good
```

```{r residuals pit.uniform}
plot(comm.nb, res.type="pit.uniform") # residuals vs fitted, pretty good
```

```{r anova manyglm}
community_now_mva.anova <- anova(comm.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=10, show.time="all")  # nBoot=10 requires ~2.3 mins, time scales linearly with nBoot
```

```{r mvabund anova p-values table}
community_now_mva.anova$table
```

The next thing is to compare all habitats against NF (Natural Forest). To do this, we need to make a bunch of pairwise community_now and habitat_now datasets (here, 5) and do anovas. Afterwards, we make a vector of all the p-values and correct them for the 5 tests that we did.

```{r}
nboot <- 10
```

```{r pairwise mvabund NF vs MF}
habitat_now_NF_MF <- habitat_now %>% filter(Habitat %in% c("NF", "MF"))
community_now_NF_MF <- community_now %>% filter(habitat_now$Habitat %in%  c("NF", "MF"))
community_now_NF_MF <- community_now_NF_MF %>% select_if(specnumber(community_now_NF_MF, MARGIN = 2) > 0) # keep only OTUs that appear in more than one site.
# community_now_NF_MF <- community_now_NF_MF[, colSums(community_now_NF_MF) >= 50] # alternative filtering:  remove OTUs that have fewer than, here, 50 reads, which keeps singleton OTUs in
community_now_NF_MF_mva <- mvabund(community_now_NF_MF)
is.mvabund(community_now_NF_MF_mva)
comm_NF_MF.nb <- manyglm(community_now_NF_MF_mva ~ habitat_now_NF_MF$Habitat, family="negative.binomial")
plot(comm_NF_MF.nb, res.type="pit.norm")
community_now_NF_MF_mva.anova <- anova(comm_NF_MF.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=nboot, show.time="all") 
community_now_NF_MF_mva.anova$table
```

```{r pairwise mvabund NF vs BB}
habitat_now_NF_BB <- habitat_now %>% filter(Habitat %in% c("NF", "BB"))
community_now_NF_BB <- community_now %>% filter(habitat_now$Habitat %in%  c("NF", "BB"))
community_now_NF_BB <- community_now_NF_BB %>% select_if(specnumber(community_now_NF_BB, MARGIN = 2) > 0) # keep only OTUs that appear in more than one site.
# community_now_NF_BB <- community_now_NF_BB[, colSums(community_now_NF_BB) >= 50] # alternative filtering:  remove OTUs that have fewer than, here, 50 reads, which keeps singleton OTUs in
community_now_NF_BB_mva <- mvabund(community_now_NF_BB)
is.mvabund(community_now_NF_BB_mva)
comm_NF_BB.nb <- manyglm(community_now_NF_BB_mva ~ habitat_now_NF_BB$Habitat, family="negative.binomial")
plot(comm_NF_BB.nb, res.type="pit.norm")
community_now_NF_BB_mva.anova <- anova(comm_NF_BB.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=nboot, show.time="all") 
community_now_NF_BB_mva.anova$table
```

```{r pairwise mvabund NF vs JC}
habitat_now_NF_JC <- habitat_now %>% filter(Habitat %in% c("NF", "JC"))
community_now_NF_JC <- community_now %>% filter(habitat_now$Habitat %in%  c("NF", "JC"))
community_now_NF_JC <- community_now_NF_JC %>% select_if(specnumber(community_now_NF_JC, MARGIN = 2) > 0) # keep only OTUs that appear in more than one site.
# community_now_NF_JC <- community_now_NF_JC[, colSums(community_now_NF_JC) >= 50] # alternative filtering:  remove OTUs that have fewer than, here, 50 reads, which keeps singleton OTUs in
community_now_NF_JC_mva <- mvabund(community_now_NF_JC)
is.mvabund(community_now_NF_JC_mva)
comm_NF_JC.nb <- manyglm(community_now_NF_JC_mva ~ habitat_now_NF_JC$Habitat, family="negative.binomial")
plot(comm_NF_JC.nb, res.type="pit.norm")
community_now_NF_JC_mva.anova <- anova(comm_NF_JC.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=nboot, show.time="all") 
community_now_NF_JC_mva.anova$table
```

```{r pairwise mvabund NF vs EC}
habitat_now_NF_EC <- habitat_now %>% filter(Habitat %in% c("NF", "EC"))
community_now_NF_EC <- community_now %>% filter(habitat_now$Habitat %in%  c("NF", "EC"))
community_now_NF_EC <- community_now_NF_EC %>% select_if(specnumber(community_now_NF_EC, MARGIN = 2) > 0) # keep only OTUs that appear in more than one site.
# community_now_NF_EC <- community_now_NF_EC[, colSums(community_now_NF_EC) >= 50] # alternative filtering:  remove OTUs that have fewer than, here, 50 reads, which keeps singleton OTUs in
community_now_NF_EC_mva <- mvabund(community_now_NF_EC)
is.mvabund(community_now_NF_EC_mva)
comm_NF_EC.nb <- manyglm(community_now_NF_EC_mva ~ habitat_now_NF_EC$Habitat, family="negative.binomial")
plot(comm_NF_EC.nb, res.type="pit.norm")
community_now_NF_EC_mva.anova <- anova(comm_NF_EC.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=nboot, show.time="all") 
community_now_NF_EC_mva.anova$table
```

```{r pairwise mvabund NF vs CL}
habitat_now_NF_CL <- habitat_now %>% filter(Habitat %in% c("NF", "CL"))
community_now_NF_CL <- community_now %>% filter(habitat_now$Habitat %in%  c("NF", "CL"))
community_now_NF_CL <- community_now_NF_CL %>% select_if(specnumber(community_now_NF_CL, MARGIN = 2) > 0) # keep only OTUs that appear in more than one site.
# community_now_NF_CL <- community_now_NF_CL[, colSums(community_now_NF_CL) >= 50] # alternative filtering:  remove OTUs that have fewer than, here, 50 reads, which keeps singleton OTUs in
community_now_NF_CL_mva <- mvabund(community_now_NF_CL)
is.mvabund(community_now_NF_CL_mva)
comm_NF_CL.nb <- manyglm(community_now_NF_CL_mva ~ habitat_now_NF_CL$Habitat, family="negative.binomial")
plot(comm_NF_CL.nb, res.type="pit.norm")
community_now_NF_CL_mva.anova <- anova(comm_NF_CL.nb, p.uni="adjusted", test="wald", resamp="pit.trap", nBoot=nboot, show.time="all") 
community_now_NF_CL_mva.anova$table
```

```{r pvalues list}
pvalues <- c(community_now_NF_MF_mva.anova$table[2,4], community_now_NF_BB_mva.anova$table[2,4], community_now_NF_JC_mva.anova$table[2,4], community_now_NF_EC_mva.anova$table[2,4], community_now_NF_CL_mva.anova$table[2,4])
pvalues
```

```{r fdr correction}
pvalues.corr.fdr<-p.adjust(pvalues, method = "fdr", n = length(pvalues))
pvalues.corr.fdr
```

#### vegan::adonis

```{r}
comm.ads <- adonis(community_now ~ habitat_now$Habitat, method = "jaccard", perm = 999)
comm.ads
```

```{r}
comm_NF_BB.ads <- adonis(community_now_NF_BB ~ habitat_now_NF_BB$Habitat, method = "jaccard", perm = 999)
comm_NF_BB.ads
```


#### Breakaway analysis

```{r breakaway}
####################### alpha diversity: breakaway and vegan ##

# ### XXXX use original otu table #
# library("breakaway")
# row.names(commNosr)
# BB <- commNosr[1:10,]
# CL <- commNosr[11:26,]
# EC <- commNosr[27:34,]
# JC <- commNosr[35:46,]
# MF <- commNosr[47:55,]
# NF <- commNosr[56:70,]
# 
# BB <- BB[,colSums(BB)>0] # keep those observed otu
# CL <- CL[,colSums(CL)>0]
# EC <- EC[,colSums(EC)>0]
# JC <- JC[,colSums(JC)>0]
# MF <- MF[,colSums(MF)>0]
# NF <- NF[,colSums(NF)>0]
# 
# ### breakaway ##
# BB.pa <- BB
# BB.pa[BB.pa>1] <- 1
# BB.fq <- matrix(c(colSums(BB.pa),colSums(BB)),ncol = 2)
# BB.fqb <- matrix(c(colSums(BB.pa),colSums(BB.pa)),ncol = 2)
# 
# breakaway(BB.fq)
# chao1(BB.fq)
# chao1_bc(BB.fq)
# chao_bunge(BB.fq)
# wlrm_transformed(BB.fq)
# wlrm_untransformed(BB.fq)
# 
# ###### vegan ###
# library(vegan)
# attach(habitat)
# pool1 <- specpool(commNosr,Habitat)
# pool1
# #op <- par(mfrow = c(1,2))
# boxplot(specnumber(commNosr)~Habitat)
# boxplot(specnumber(commNosr)/specpool2vect(pool1)~Habitat)
# poolaccum(commNosr)
# estaccumR(commNosr)
# 
# ### XXXX use commNosr without single read ##
# commNosr.NS <- commNosr
# commNosr.NS[commNosr.NS==1] <- 0 # drop those single read cell
# sum(commNosr) - sum(commNosr.NS)
# 
# BB.NS <- commNosr.NS[1:10,]
# CL.NS <- commNosr.NS[11:26,]
# EC.NS <- commNosr.NS[27:34,]
# JC.NS <- commNosr.NS[35:46,]
# MF.NS <- commNosr.NS[47:55,]
# NF.NS <- commNosr.NS[56:70,]
# # 
# BB.NS <- BB.NS[,colSums(BB.NS)>0] 
# CL.NS <- CL.NS[,colSums(CL.NS)>0]
# EC.NS <- EC.NS[,colSums(EC.NS)>0]
# JC.NS <- JC.NS[,colSums(JC.NS)>0]
# MF.NS <- MF.NS[,colSums(MF.NS)>0]
# NF.NS <- NF.NS[,colSums(NF.NS)>0]
# 
# ##
# 
# BB.NS.pa <- BB.NS
# BB.NS.pa[BB.NS.pa>1] <- 1
# BB.NS.fq <- matrix(c(colSums(BB.NS.pa),colSums(BB.NS)),ncol = 2)
# BB.NS.fqb <- matrix(c(colSums(BB.NS.pa),colSums(BB.NS.pa)),ncol = 2)
# 
# breakaway(BB.NS.fq)
# breakaway(BB.NS.fqb)
# chao1(BB.NS.fq)
# chao1_bc(BB.NS.fq)
# chao_bunge(BB.NS.fq)
# wlrm_transformed(BB.fq)
# wlrm_untransformed(BB.NS.fq)
# 
# ###### vegan ###
# #library(vegan)
# #attach(habitat)
# pool2 <- specpool(commNosr.NS,Habitat)
# pool2
# #op <- par(mfrow = c(1,1))
# boxplot(specnumber(commNosr.NS)~Habitat)
# boxplot(specnumber(commNosr.NS)/specpool2vect(pool2)~Habitat)
# poolaccum(commNosr.NS)
# estaccumR(commNosr.NS)
# 
# ### XXXX otu table with log2(reads number), those single read also dropped ##
# library("breakaway")
# commNosr.log <- commNosr
# commNosr.log <- log2(commNosr.log[1:70,]+0.1) 
# commNosr.log[commNosr.log<0] <- 0 
# commNosr.log <- floor(commNosr.log[1:70,]) #
# 
# row.names(commNosr.log)
# BB <- commNosr.log[1:10,]
# CL <- commNosr.log[11:26,]
# EC <- commNosr.log[27:34,]
# JC <- commNosr.log[35:46,]
# MF <- commNosr.log[47:55,]
# NF <- commNosr.log[56:70,]
# 
# BB <- BB[,colSums(BB)>0] # keep those observed otu
# CL <- CL[,colSums(CL)>0]
# EC <- EC[,colSums(EC)>0]
# JC <- JC[,colSums(JC)>0]
# MF <- MF[,colSums(MF)>0]
# NF <- NF[,colSums(NF)>0]
# 
# ### breakaway ##
# BB.pa <- BB
# BB.pa[BB.pa>1] <- 1
# BB.fq <- matrix(c(colSums(BB.pa),colSums(BB)),ncol = 2)
# BB.fqb <- matrix(c(colSums(BB.pa),colSums(BB.pa)),ncol = 2)
# 
# breakaway(BB.fq)
# chao1(BB.fq)
# chao1_bc(BB.fq)
# chao_bunge(BB.fq)
# wlrm_transformed(BB.fq)
# wlrm_untransformed(BB.fq)
# 
# ###### vegan ###
# library(vegan)
# #attach(habitat)
# pool1 <- specpool(commNosr.log,Habitat)
# pool1
# #op <- par(mfrow = c(1,2))
# boxplot(specnumber(commNosr.log)~Habitat)
# boxplot(specnumber(commNosr.log)/specpool2vect(pool1)~Habitat)
# poolaccum(commNosr.log)
# estaccumR(commNosr.log)
# ##

###### use binary otu table without single read  #######

gfgMB <- read.table("./data/2libs_otu_WithTax4R_orig.txt", header = TRUE, na.strings = "NA")
habitat <- select(gfgMB, Site:Weather)
community <- select(gfgMB, starts_with("Cluster_"))

# habitat <- gfgMB[, 1:5]
# community <- gfgMB[,10:4973]
commNosr <- community[, colSums(community) >= 50] ## change to this data 

commNosrB <- commNosr
commNosrB[commNosrB==1] <- 0 # drop those single read cell
sum(commNosr) - sum(commNosrB)
commNosrB[commNosrB>1] <- 1

## as breakaway required, make a frequency table for each habitat
habitat_now$Site
# row.names(commNosrB)


BB <- commNosrB[1:10,]
CL <- commNosrB[11:26,]
EC <- commNosrB[27:34,]
JC <- commNosrB[35:46,]
MF <- commNosrB[47:55,]
NF <- commNosrB[56:70,]

BB <- BB[,colSums(BB)>0] # keep those observed otu 
CL <- CL[,colSums(CL)>0]
EC <- EC[,colSums(EC)>0]
JC <- JC[,colSums(JC)>0]
MF <- MF[,colSums(MF)>0]
NF <- NF[,colSums(NF)>0]

####

BB.pa <- BB
BB.pa[BB.pa>1] <- 1
BB.fq <- matrix(c(colSums(BB.pa),colSums(BB)),ncol = 2)
BB.fqb <- matrix(c(colSums(BB.pa),colSums(BB.pa)),ncol = 2)


breakaway(BB.fq)
chao1(BB.fq)
chao1_bc(BB.fq)
chao_bunge(BB.fq)
wlrm_transformed(BB.fq)
wlrm_untransformed(BB.fq)

###### vegan ###
library(vegan)
attach(habitat)
pool1 <- specpool(commNosrB, habitat$Habitat)
pool1
#op <- par(mfrow = c(1,2))
boxplot(specnumber(commNosrB)~Habitat)
boxplot(specnumber(commNosrB)/specpool2vect(pool1)~Habitat)
poolaccum(commNosrB)
estaccumR(commNosrB)

############## Boral #######
##
#mcmc.control <- list(n.burnin = 100, n.iteration = 1000, n.thin = 10)
mcmc.control <- list(n.burnin = 300, n.iteration = 3000, n.thin = 30)
colnames(commNosrB) <- c(1:520)
comm.fit.p2 <- boral(commNosrB, family = "poisson", num.lv = 2, row.eff = "fixed", calc.ics = FALSE, mcmc.control = mcmc.control)
summary(comm.fit.p3)
par(mfrow = c(2,2))
plot(comm.fit.p3) ## Plots used in residual analysis, 
#dev.off()
par(mfrow = c(1,1))
lvsplot(comm.fit.p2, biplot = FALSE, jitter = TRUE)


```

```{r}
comm.fit.p2d <- boral(community, family = "poisson", num.lv = 2, row.eff = "fixed", calc.ics = FALSE)
summary(comm.fit.p2d)
par(mfrow = c(2,2))
plot(comm.fit.p2d) ## Plots used in residual analysis, 
#dev.off()
par(mfrow = c(1,1))
lvsplot(comm.fit.p2d)

comm.fit.p3d <- boral(commNosrB, family = "poisson", num.lv = 3, row.eff = "fixed", calc.ics = FALSE)
summary(comm.fit.p3d)
par(mfrow = c(2,2))
plot(comm.fit.p3d) ## Plots used in residual analysis, 
#dev.off()
par(mfrow = c(1,1))
lvsplot(comm.fit.p3d)

#comm.fit.p4d <- boral(commNosrB, family = "poisson", num.lv = 4, #row.eff = "fixed", calc.ics = FALSE)
#summary(comm.fit.p4d)
#par(mfrow = c(2,2))
#plot(comm.fit.p4d) ## Plots used in residual analysis, 
##dev.off()
#par(mfrow = c(1,1))
#lvsplot(comm.fit.p4d)


x <- habitat$Habitat
commB.fit <- boral(commNosrB, X = x, family = "poisson", num.lv = 0, mcmc.control = mcmc.control) # boral fit failed...

#####
vegemite(commNosrB)
tabasco(commNosrB)
x0 <- colSums(commNosrB[1:10,])
x1 <- colSums(commNosrB[11:26,])
x2 <- colSums(commNosrB[27:34,])
x3 <- colSums(commNosrB[35:46,])
x4 <- colSums(commNosrB[47:55,])
x5 <- colSums(commNosrB[56:70,])

habSum <- matrix(c(x0,x1,x2,x3,x4,x5),ncol = 520)
colnames(habSum)
rownames(habSum)
row <- c("BB","CL","EC","JC","MF","NF")
col <- c(1:520)
dimnames(habSum) <- list(row,col)

vegemite(habSum,scale = "Hill")
tabasco(habSum)

habSum1 <- matrix(c(x0,x2,x3,x4,x5),ncol = 520)
row <- c("BB","EC","JC","MF","NF")
col <- c(1:520)
dimnames(habSum1) <- list(row,col)

habSum1 <- habSum1[,colSums(habSum1)]

vegemite(habSum1,scale = "Hill")
tabasco(habSum1)



#c,ommBT <- t(commNosrB)
#row.names(commBT)
#x <- rowSums(commBT[,1:10])
#x1 <- rowSums(commBT[,11:26])
#x2 <- rowSums(commBT[,27:34])
#x3 <- rowSums(commBT[,35:46])
#x4 <- rowSums(commBT[,47:55])
#x5 <- rowSums(commBT[,56:70])
#
#a <- matrix(c(x,x1,x2,x3,x4,x5),ncol = 6)
#
#vegemite(a)
#tabasco(a)

###### ####

```

```{r } 
######### get original data and do filter
gfgMB <- read.table("data/2libs_otu_WithTax4R.txt",header=TRUE, sep="\t", row.names=1, na.strings="NA", dec=".", strip.white=TRUE) #this otu table without taxon-info
row.names(gfgMB)
#colnames(gfgMB)
habitat <- gfgMB[1:70,1:4]
community <- gfgMB[1:70,5:4968]
taxinfo <- gfgMB[71:77,5:4968]

write.csv(community,'SCMBC_otu.csv')
community <- read.csv('SCMBC_otu.csv',row.names=1, header=T)
commNosr <- community[,colSums(community)>=50]
colnames(commNosr)

Tax <- taxinfo[,c(colnames(commNosr))]
Tax <- t(Tax)
#comTax <- 


######### alpha diversity  ############
## by specpool
library(vegan)
commNosrB <- commNosr
commNosrB[commNosrB<3] <- 0 # according before results and field work experience, JC forest seems get more alpha diversity. so drop those single read and double reads cell
commNosrB[commNosrB>1] <- 1

attach(habitat)
pool1 <- specpool(commNosrB,Habitat)
pool1 ## seems better
#	Species	chao	chao.se	jack1	jack1.se	jack2	boot	boot.se	n
#BB	80	196.4462	45.76738	132.2000	20.30714	169.3556	101.70195	9.500782	10
#CL	180	316.4667	39.00446	269.0625	29.35338	326.8708	218.43714	14.985708	16
#EC	62	110.5962	21.88252	95.2500	15.68738	115.3929	76.53721	8.226332	8
#JC	90	225.5256	52.28162	146.8333	22.26576	188.6515	113.59180	10.595451	12
#MF	118	368.6944	82.61292	202.4444	33.71046	265.4444	152.69311	14.433310	9
#NF	195	457.2968	67.57570	318.2000	42.14641	407.6524	246.30169	20.212816	15

####  try breakaway ###
BB <- commNosrB[1:10,]
CL <- commNosrB[11:26,]
EC <- commNosrB[27:34,]
JC <- commNosrB[35:46,]
MF <- commNosrB[47:55,]
NF <- commNosrB[56:70,]

BB <- BB[,colSums(BB)>0] # keep those observed otu 
CL <- CL[,colSums(CL)>0]
EC <- EC[,colSums(EC)>0]
JC <- JC[,colSums(JC)>0]
MF <- MF[,colSums(MF)>0]
NF <- NF[,colSums(NF)>0]

BB.fq <- matrix(c(colSums(BB),colSums(BB)),ncol = 2)
#BB.fq <- BB.fq[sort(BB.fq[1,]),]
BB.fq <- BB.fq[order(rowSums(BB.fq)),]
BB.fq.ns <- BB.fq[rowSums(BB.fq)>2,]

library("breakaway")
breakaway(BB.fq)
chao1(BB.fq)
chao1_bc(BB.fq)
chao_bunge(BB.fq)
wlrm_transformed(BB.fq)
wlrm_untransformed(BB.fq)
breakaway_nof1(BB.fq.ns)
#
CL.fq <- matrix(c(colSums(CL),colSums(CL)),ncol = 2)
CL.fq <- CL.fq[order(rowSums(CL.fq)),]
CL.fq.ns <- CL.fq[rowSums(CL.fq)>2,]
#CL.fq.ns
breakaway(CL.fq)
breakaway_nof1(CL.fq.ns)
#
EC.fq <- matrix(c(colSums(EC),colSums(EC)),ncol = 2)
EC.fq <- EC.fq[order(rowSums(EC.fq)),]
EC.fq.ns <- EC.fq[rowSums(EC.fq)>2,]
breakaway(EC.fq)
breakaway_nof1(EC.fq.ns)
#
JC.fq <- matrix(c(colSums(JC),colSums(JC)),ncol = 2)
JC.fq <- JC.fq[order(rowSums(JC.fq)),]
JC.fq.ns <- JC.fq[rowSums(JC.fq)>2,]
breakaway(JC.fq)
breakaway_nof1(JC.fq.ns)
#
MF.fq <- matrix(c(colSums(MF),colSums(MF)),ncol = 2)
MF.fq <- MF.fq[order(rowSums(MF.fq)),]
MF.fq.ns <- MF.fq[rowSums(MF.fq)>2,]
breakaway(MF.fq)
breakaway_nof1(MF.fq.ns)
#
NF.fq <- matrix(c(colSums(NF),colSums(NF)),ncol = 2)
NF.fq <- NF.fq[order(rowSums(NF.fq)),]
NF.fq.ns <- NF.fq[rowSums(NF.fq)>2,]
breakaway(NF.fq)
breakaway_nof1(NF.fq.ns)


### original data without singletons and doubletons: alpha diversity
# by specpool
library(vegan)
commNosrN <- commNosr
commNosrN[commNosrN<3] <- 0 # according before results and field work experience, JC forest seems get more alpha diversity. so drop those single read and double reads cell
#commNosrB[commNosrB>1] <- 1

attach(habitat)
pool2 <- specpool(commNosrN,Habitat)
pool2 ## seems better

### try breakaway ##
BBN <- commNosrN[1:10,]
CLN <- commNosrN[11:26,]
ECN <- commNosrN[27:34,]
JCN <- commNosrN[35:46,]
MFN <- commNosrN[47:55,]
NFN <- commNosrN[56:70,]

BBN <- BBN[,colSums(BBN)>0] # keep those observed otu 
CLN <- CLN[,colSums(CLN)>0]
ECN <- ECN[,colSums(ECN)>0]
JCN <- JCN[,colSums(JCN)>0]
MFN <- MFN[,colSums(MFN)>0]
NFN <- NFN[,colSums(NFN)>0]

BBN.pa <- BBN
BBN.pa[BBN.pa>1] <- 1
BBN.fq <- matrix(c(colSums(BBN.pa),colSums(BBN)),ncol = 2)

#BB.fq <- matrix(c(colSums(BB),colSums(BB)),ncol = 2)
#BB.fq <- BB.fq[sort(BB.fq[1,]),]
BBN.fq <- BBN.fq[order(rowSums(BBN.fq)),]
BBN.fq.ns <- BBN.fq[BBN.fq>1]


library("breakaway")
breakaway(BBN.fq)
#


########### phylocurve
install.packages("phylocurve")
install.packages("betapart")


###### beta diversity ######
### PCA 
install.packages("SparseM") ## need by library 'car'
library(foreign)
library(vegan)
library(car)
commNosrB <- commNosr
commNosrB[commNosrB<3] <- 0 # according before results and field work experience, JC forest seems get more alpha diversity. so drop those single read and double reads cell
commNosrB[commNosrB>1] <- 1

hab <- habitat[,2:4]
comm.pca <- rda(commNosrB)
comm.pca <- rda(commNosrB~Habitat,habitat)
comm.pca <- rda(commNosrB~Habitat*Altitude*Weather,habitat)
summary(comm.pca)
plot(comm.pca)



#rowSums(commNosrB) # JC04 has no otu


#### boral
library("boral")
library(mvabund)



# #######


```

